### 2016-10-31 answer_to_referees.txt
More modifications + answer to referres

Answer to Reviewer 1
====================

* Suggest also citing http://nsls-ii.github.io/ and                  
http://cars9.uchicago.edu/software/python/pyepics3/ along with pytango.

We have cited PyEPICS as well as NSLS-II.

* Figure 3 should use the mpl OO API instead of plt.  A gist does not
  seem like a sufficiently stable citation target.

plt is used in a lot of code examples for legacy reasons, but the OO API
of matplotlib is indeed preferable nowadays. We have modified Figure 3
accordingly. We generated a DOI with Zenodo for a more stable citation
target.

* The discussion about skimage functionality (fig 5) was very fast and
  felt a bit disjointed.

Image processing is a very broad field, and scikit-image provides a large
variety of processing operations. Therefore, we chose to illustrate only
a few capabilities of scikit-image, since a lengthy list would bore the
reader, while not being very instructive. Nevertheless, the choice of
these operations, and the link between them, might have been better
explained. We now explain why denoising, object detection and
measurements are important for processing microtomography images "In the
following, we briefly illustrate how scikit-image can be used for some
typical image processing tasks encountered when analyzing tomographic
images: denoising, mid-range feature detection, segmentation and
measurement of region properties. For the sake of brevity, other tasks
such as contrast manipulation or geometric transformations are not
described here; the interested reader is referred to the documentation
of scikit-image.", "Denoising data is often the first step of an image
processing workflow", "Detecting the presence of objects or extracting
pixels corresponding to objects (a task known as segmentation) is an
important task of image analysis for medical or materials science
applications.". We have also improved transitions between paragraphs.

* Given the recent discussions of system package managers vs conda/EPD vs
  pip (ex
http://www.curiousefficiency.org/posts/2016/09/python-packaging-ecosystem.html)
I suggest re-wording the getting started section.

Python packaging is a broad topic, that goes well beyond the scope of our
paper. As explained in the blog post cited by the referee, different
solutions each have their pros and cons. The purpose of the getting
started section is mostly to give some pointers to the readers, rather
than engaging in a discussion on packaging. For the sake of completeness,
we now mention conda as a command-line tool for Anaconda (the most common
use of conda).

Answer to Reviewer 2
====================

* The article "Analyzing X-ray images in Python with scikit-image"
describes a powerful open software package which is of high use to
analyze images. It is well written, of high importance for all people
dealing with heavy data-image processing. Hence, I can recommend
publication.

We thank the referee for his/her positive comments about our paper.

* the title is a bit misleading as it explicitly states "X-ray images"
  while the routines described are quite general, i.e.
would also work for MRI or confocal laser scanning data sets, indeed,
there is  no hard link to X-ray. Contrary, X-ray images
are a broader term then just the microtomographic data included in this
article: fluorescence mapping reveals X-ray images,
each pixel contains a spectrum (hyperspectral images) - something not to
be     considered with scikit but the mentioned PyMCA,
i.e. the latter is a dedicated X-ray software, the same for xrdua and
diffraction data / tomography. Especially as the journal
title contains "Chemical Imaging" this needs to be more clear considering
that  standard full-field X-ray imaging is a non-
analytical tool. The referee is not the most creative person in the
world, a    suggestion for a modification could look like that
"Analyzing image data in Python with scikit-image: the case of
microtomography" or similar

We understand the concern of the referee, and we have changed the title
to "Analyzing microtomography data with Python and the scikit-image
library".

- p. 1, l. 51 / right column: the list is a bit incomplete, as stated above,    
xrdua could be mentioned, also in terms of volume                               
image processing there is powerful commercial and non-commercial software we    
should be listed here as a matter of fairness so                                
the reader can make his own choice, i.e. AVS www.avs.com, pore3D from Elettra   
Synchrotrone, ToolIP

In addition to Aviso Fire, we now mention pore3D and ToolIP/Mavi. We also
mention XRDUA for processing powder diffraction images. To our knowledge,
AVS was more popular in the past that it is now in the tomography
community, therefore we preferred to cite the most popular software only.

- the Figures appear not ordered in the manuscript, i.e. Fig 1, Fig 4, Fig 2    
....

We apologize for this inconvenience, that is due to Latex optimizing
float placement. This should be corrected in the publisher's layout.

- p. 5, l. 47, left: "diffraction images" - define what precisely is considered,
i.e. Bragg diffraction images?

We now say "peaks in 2D Bragg diffraction patterns".

- Fig 3 and Fig 6 the font size in the images itself are on the limit,
  make sure they can still be read in the final article

We increased the font size in Figures 3 and 6 (modifying the layout of
Figure 6 to this end).

- Fig 5: for the non-synthetic data sets: could one perhaps state where the data
was taken, i.e. in the acknowledgement? Scale                                   
bars are also missing.

Scale bars were actually omitted on purpose, since Fig. 5 is a very busy
figure, and its interest lies mostly in the application of image
processing operations, not in the original samples and images
themselves.

We have acknowledged the origin of the membrane image of Fig. 5 (b),
which was the only non-synthetic image for which the origin was missing.

- p. 6, l. 47 and later: for example for the Watershed it is not only           
interesting to know the raw CPU consumption, also the                           
memory consumption is important: at least in the early days a watershed would   
need 16x[image size] as amount of memory to                                     
process which was often more limiting than the CPU time, same for the euclidean 
distance transformation

The watershed algorithm implemented in scikit-image (from Soille,
"Automated Basin Delineation from Digital Elevation Models Using
Mathematical Morphology", Signal Processing 20 (1990) 171-182.) is much
more memory-savvy, since it uses of the order of 5x as much memory as the
original image size. Indeed, the algorithm does not need a lot of
additional variables of the same size of the original image.

Although scikit-image is not designed specifically for images of the size
of tomography volumes, memory consumption is often improved and optimized
during the code review process. We now write: "During the code review
process, a close watch is also kept on memory consumption, since for
large image sizes, transfers between computer memory (RAM) and CPU cache
are often a serious performance bottleneck."

- Ref Armando Sole: perhaps consider also citing the corresponding paper        
V.A. Solé, E. Papillon, M. Cotte, Ph. Walter, J. Susini, A multiplatform code   
for the analysis of energy-dispersive X-ray fluorescence spectra, Spectrochim.  
Acta Part B 62 (2007) 63-68.

We are now citing this paper by Solé et al.